XOR-Some Prolific Pilot 3

### 2021-07-25 analysis/xor-some-prolific-pilot3.Rmd
analyse pilot 3 with updated vignettes

---
title: "XOR-Some Prolific Pilot 2"
author: "Polina Tsvilodub"
date: "4/2/2021"
output: github_document
---

In this second pilot on Prolific we gather a more substantial amount of pilot data for the xor-some study on Prolific. The structure of the experiment was as follows:
Participants read instructions, completed three example trials, and then completed 8 main blocks consisting of 4 xor and 4 some items. Each main block had the following structure: Participants read the background story, answered one comprehension question, then answered competence / relevance / prior questions in randomized order; then they read another 3 comprehension questions, after which the critical utterance was added below the background story. They answered the inference strength question, and then competence / relevance questions in randomized order again.

N=118 participants were recruited for this pilot and compensated 2 pounds/participant. 8 items (4 some + 4 xor) were sampled at random for each participant such that each participant saw one item in each condition (relevance X competence X prior = 8 unique conditions).

@@

```{r anonymization, include=FALSE}
#d_raw <- read_csv("~/Documents/Research/XOR/raw_data/results_58_xor-some-Prolific-pilot2_N120.csv")
# remove IDs
#d_raw %>% select(-prolific_id) %>% write_csv("../data/pilots/results_58_xor-some-Prolific-pilot2_N120.csv") 
```

## Checks & Exclusions
```{r data, results='hide', include=FALSE}
d <- read_csv("../data/pilots/results_58_xor-some-Prolific-pilot2_N120.csv")
```

@@

Check native languages. Participants not indicating English as (one of) their native language(s) are be excluded. We also compute some participant demographics.
```{r languages}
d %>% distinct(languages)
cat("Number of partiipants before excluding non-natives: ", d %>% distinct(submission_id) %>% count() %>% pull() ) 
# exclude non-natives if necessary
d_native <- d %>% 
 # filter(("en" | "En" | "bri") %in% languages)
  filter(grepl("[(en)(br)]", languages, ignore.case = T))

cat("Number of partiipants after excluding non-natives: ", d_native %>% distinct(submission_id) %>% count() %>% pull() )

cat("Mean age: ", d_native %>% pull(age) %>% mean(., na.rm = T) )
d_native %>% count(gender) %>% mutate(n = n/88)
```

@@

```{r counts}
# check xor/some vs. trial type
d_native %>% count(main_type, condition) 
# check xor/some vs. experimental condition
d_native %>% count(main_type, exp_condition)
# count items used
d_native %>% count(title)
```

@@

Investigate very fast completion times:
```{r}
 d_native %>% mutate(timeSpent = round(timeSpent, 2)) %>%
  filter(timeSpent < 8) -> fast_completions

# check if their responses look somehow peculiar
fast_completions %>%
  ggplot(., aes(x = response)) +
  geom_histogram()

@@

# get participants failing comprehension questions
d_test <- d_test %>%
  group_by(submission_id) %>%
  mutate(passed_filler_trial = case_when(test_condition == "true" ~ response >= 70,
                                   test_condition == "false" ~ response <= 40,
                                   test_condition == "uncertain" ~ response %in% (0:90)),
         mean_comprehension = mean(passed_filler_trial),
         passed_filler = mean_comprehension >= 0.8
         )

@@

Check exclusions if applying exclusion ranges as suggested by Bob -- then almost twice as many participants are excluded based on their comprehension questions performance (I believe mainly due to a lower bound on correct responses in the uncertain questions). 
```{r}
# true comprehension questions: 50-100
# uncertain comprehension questions: 25-75
# false comprehension questions: 0-50
# get participants failing example trials

d_exmpl2 <- d_exmpl %>% group_by(submission_id) %>% 
  mutate(example_condition = ifelse(grepl("as certainly true", question), "true", 
                                    ifelse(grepl("as certainly false", question), "false",
                                           "uncertain")),
         passed_example_trial = case_when(example_condition == "true" ~ response >= 50,
                                    example_condition == "false" ~ response <= 50,
                                    example_condition == "uncertain" ~ response %in% (25:75)),
       #  check if all trials passed
         passed_example = case_when(sum(passed_example_trial) == 0 ~ FALSE,
                                    TRUE ~ TRUE)
         ) %>% filter(passed_example == F)

cat("Subjects failing the example trials: ", d_exmpl2 %>% distinct(submission_id) %>% pull() %>% length())

# get participants failing comprehension questions
d_test2 <- d_test %>%
  group_by(submission_id) %>%
  mutate(passed_filler_trial = case_when(test_condition == "true" ~ response >= 50,
                                   test_condition == "false" ~ response <= 50,
                                   test_condition == "uncertain" ~ response %in% (25:75)),
         mean_comprehension = mean(passed_filler_trial),
         passed_filler = mean_comprehension >= 0.8
         )

d_test_fail2 <- d_test2 %>% 
  filter(passed_filler == F)

cat(" Subjects failing the comprehension trials: ", d_test_fail2 %>% distinct(submission_id) %>% pull() %>% length())

```
Now we explore applying exclusion criteria on z-scored data. Here it might make sense to exclude participants whose >=0.8 of responses are more than 2 SD from the mean (?). 
```{r}
# group by extended block, exmpl / filler / critical, maybe polarity of filler exmpl question
# add block extension, put all data together somehow
d_critical_fail_zScored <- d_main %>% 
  mutate(block_extended = ifelse(
    !(ifelse(is.na(critical_question), F, T)), 
    block, 
    ifelse(block %in% c("some", "xor"), "target", str_c(block, "_wUtt", ""))
  )  
  ) %>% 
  filter(!(grepl("test", block_extended))) %>%
  group_by(submission_id, block_extended) %>%
  mutate(subj_block_mean = mean(response),
         subj_block_sd = sd(response),
         response_centered = (response - subj_block_mean)/subj_block_sd,
         # catch the cases where sd is 0 
         response_centered = ifelse(is.na(response_centered), 0, response_centered)) %>%
  ungroup() %>%
  # add by-block SDs, such that participants can be excluded based on being outside 2 SDs on a block
  # or should this be based on by-subject SDs???? 
  group_by(block_extended) %>%
  mutate(block_mean = mean(response),
         block_sd = sd(response),
         passed_main_trial = case_when(abs(response_centered) <= 2*block_sd ~ T,
                                       T ~ F),
         mean_critical_passRate = mean(passed_main_trial))

# get participants failing comprehension questions with z-scored data; here, by-subject SD are used for the exclusion
d_test_fail_zScore <- d_test %>%
  group_by(submission_id, test_condition) %>%
  mutate(
    subj_block_mean = mean(response),
    subj_block_sd = sd(response),
    response_centered = (response - subj_block_mean)/subj_block_sd,
    response_centered = ifelse(is.na(response_centered), 0, response_centered),
    passed_filler_trial = case_when(abs(response_centered) <= 2*subj_block_sd ~ T,
                                   T ~ F),
         mean_comprehension = mean(passed_filler_trial),
         passed_filler = mean_comprehension >= 0.8
         )

d_test_fail_zScore <- d_test_fail_zScore %>% 
  ungroup() %>%
  filter(passed_filler == F)

cat(" Subjects failing the comprehension trials when using z-scored raw data: ", d_test_fail_zScore %>% distinct(submission_id) %>% pull() %>% length())

@@

```{r clean}
# get overall mean ratings / subject
d_full_clean %>% group_by(submission_id) %>% summarise(mean_rating = mean(response)) %>% arrange(mean_rating)
```

@@

Create more extensive condition labels, including information about whether a rating was produced with or without the utterance given.

@@

Compute the average deviation from expected responses for each vignette, individually for each dimension (rel / comp / pri):
```{r}
d_critical_long %>% 
  filter(block != "xor" & block != "some") %>%
  filter(block == class_condition) %>%
  mutate(expected_rating = prior_class * 100,
         response_deviation = abs(response - expected_rating),
         block = ifelse(block == "xor" | block == "some", "target", block)
         ) %>%
  group_by(title, block) %>%
  summarize(mean_deviation = mean(response_deviation)) %>%
  arrange(desc(mean_deviation), .by_group = T) %>%
  # get the largest deviations - over 50 
  filter(mean_deviation >= 50) -> d_critical_deviations

#d_critical_deviations %>% write_csv("../data/pilots/pilot2_byItem_rating_deviations.csv")

@@

#d_critical_sd %>% write_csv("../data/pilots/pilot2_byItem_rating_SDs.csv")
```

Visually check the spread of by-item ratings among subjects for each question:
```{r, fig.height=25, fig.width=10}
d_critical_long_grandMeans <- d_critical_long %>% 
  mutate(title = paste(title, exp_condition, sep = "_"),
         block = ifelse(block == "xor" | block == "some", "target", block)) %>%
  filter(block == class_condition | block == "target") %>%
  group_by(main_type, block) %>% summarize(mean_resp = mean(response))

d_critical_long_expectedCats <- d_critical_long %>% 
  mutate(title = paste(title, exp_condition, sep = "_"),
         block = ifelse(block == "xor" | block == "some", "target", block)) %>%
  filter(block == class_condition | block == "target") %>%
  group_by(main_type, block, title) %>% summarize(mean_resp_block = mean(response))

d_critical_long %>% 
  mutate(title = paste(title, exp_condition, sep = "_"),
         block = ifelse(block == "xor" | block == "some", "target", block)) %>%
  left_join(., d_critical_long_grandMeans, by = c("main_type", "block")) %>% filter(block == class_condition | block == "target") -> d_long_place_holder

left_join(d_long_place_holder, d_critical_long_expectedCats, by = c("main_type", "block", "title")) %>% 
  mutate(expected_block = ifelse(block != "target", prior_class * 100, NA),
         block = factor(block, levels = c("relevance", "competence", "prior"))) %>%
  ggplot(., aes(x = block, y = response, color = block)) +
  geom_point(alpha = 0.7, position = position_jitter(width = 0.1)) +
  geom_hline(aes(yintercept = mean_resp), alpha = 0.7) +
  geom_point(aes(y = mean_resp), color = "black", shape = 2) +
  geom_point(aes(y = mean_resp_block, fill = block), color = "black") +
  geom_point(aes(y = expected_block), color="red", size = 2.5) +
  ylab("Raw responses to respective questions") +
  facet_wrap(main_type~title, ncol = 4) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ggtitle("By-item by-question raw ratings\nBlack dots indicate by-block means within-item, lines and triangles indicate grand by-block means\nRed dots indicate expected rating")
```
For the following items, participants seem  to *disagree* in their judgments for all of the predictors: Emails from a broken laptop, Gift unwrapping, Martha's cookies, Suzie's fruits, Brad's clothes, Crime, maybe more..

@@

## Plots
Plot main rel / comp / prior questions by main condition and by prior classification of the item (x-axis), separated into with / without critical utterance (shape, color). The prior questions were only used without the utterance. 
```{r}

d_critical_long %>% 
  filter(block != "xor" & block != "some") %>%
  filter(block == class_condition) %>%
  group_by(main_type, class_condition, w_utterance, prior_class) %>% 
  summarize(mean_response = mean(response)) -> d_critical_summary

d_critical_long %>% 
  filter(block != "xor" & block != "some") %>%
  filter(block == class_condition) %>%
  ggplot(., aes(x = as.factor(prior_class), y = response, shape = w_utterance, color = w_utterance)) +
  geom_point(size = 2, alpha = 0.6, position = position_jitter(width = 0.1)) +
  geom_point(data = d_critical_summary, aes(x = as.factor(prior_class), y = mean_response, shape = w_utterance), 
             color = "red", size = 3) +
  ylab("Responses to respective predictor question") +
  xlab("Anticipated categorization of the items (low / high)") +
  facet_wrap(main_type~class_condition) # get ratings from the respective trials only

```{r, fig.height=3}
# check whether the ratings in the two prior questions are the same in xor 
d_critical_long %>% 
  filter(block != "xor" & block != "some") %>%
  filter(block == class_condition) %>%
  filter(block == "prior", main_type == "xor") %>%
  mutate(priorQ_nr = rep(c(1,2), 396),
         unique_ID = paste(submission_id, ID, sep="_")) -> d_xor_priors
d_xor_priors %>% group_by(priorQ_nr, prior_class) %>%
  summarise(mean = mean(response)) -> d_xor_priors_summary

d_xor_priors %>%
  ggplot(., aes(x = as.factor(priorQ_nr), y = response )) +
  geom_point(size = 2, alpha = 0.6, position = position_jitter(width = 0.1)) +
  geom_point(data = d_xor_priors_summary, aes(x = as.factor(priorQ_nr), y = mean), color = "red", size = 3) +
  geom_path(aes(group = "unique_ID"), alpha = 0.6) +
  ylab("Responses to prior questions") +
  xlab("First vs Second prior question for high vs low prior conditions") +
  facet_wrap(~as.factor(prior_class)) + # get ratings from the respective trials only 
  ggtitle("Xor prior question ratings.\nLines indicate the two responses provided by one participant for a specific item.")
```
```{r}
# check analytically how many subjects rate prior questions differently
d_xor_priors %>% select(-trial_number, -RT, -prompt) %>% 
  pivot_wider(names_from = "priorQ_nr", values_from = "response") %>%
  mutate(response_difference = abs(`1` - `2`),
         bigDifference = ifelse(response_difference > 15, T, F)) -> d_xor_priors_differences

# 41 responses with difference > 15 between first and second rating
d_xor_priors_differences %>% count(bigDifference)

@@

Plot inference ratings as a function of *anticipated* rating of the explanatory factor, similar to the paper: 
``` {r}
d_critical_long %>% 
  filter(block == "xor" | block == "some") %>%
  ggplot(., aes(x = prior_class, y = response)) +
  geom_point(size = 2, alpha = 0.5) +
  geom_smooth(method="lm") +
  ylab("Inference strength ratings") +
  facet_wrap(block~class_condition) +
  xlab("Anticipated categorization of the items") +
  ggtitle("Inference strength ratings by-predictor")

```{r}
# make a versatile wide representation of the critical data
d_critical_wide <- d_critical_long %>% 
  select(submission_id, title, main_type, block_extended, response) %>% 
  unique() %>% 
  pivot_wider(
    names_from = block_extended, 
    values_from = response, 
    values_fn = mean # getting means for double prior measurement in "xor"
  ) 
```

Plot inference strength ratings against raw predictor ratings from each participant across items:
```{r}
d_critical_wide %>%
  ggplot(., aes(x = relevance, y = target)) +
  geom_point(size = 2, alpha = 0.7) +
  geom_smooth(method = "lm") +
  ylab("Inference strength ratings") +
  facet_wrap(~main_type, ncol = 1) -> p.rel

@@

gridExtra::grid.arrange(p.rel, p.comp, p.pri, ncol = 3) 
```

Plot mean ratings (across all participants) for each vignette (with its respective condition indicated in the facet title, in the order relevance/competence/prior), in each condition. For relevance and competence, the color indicates whether it was presented without the utterance or with. 
```{r, fig.width=10, fig.height=30}
bar.width = 0.8
d_critical_long %>% 
  mutate(title = paste(title, exp_condition, sep = "_")) %>%
  filter(block == class_condition | block == "xor" | block == "some") %>%
  group_by(block_extended, title, w_utterance, main_type) %>%
  summarize(mean_rating = mean(response)) %>% 
  ggplot(., aes(x = block_extended, y = mean_rating, fill = w_utterance)) +
  geom_col(alpha = 0.7, width = bar.width, position = position_dodge(width = bar.width)) +
  #geom_point(size = 2, alpha = 0.5, position = position_jitter(width = 0.1)) +
  ylab("Mean responses to respective questions") +
  xlab("Question type") +
  facet_wrap(main_type~title, ncol = 4) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ggtitle("By-item by-question mean ratings")
```

Plot by-item trends between predictor ratings and target inference strength ratings (relevance / competence ratings with utterance are omitted):
```{r, fig.height=35, fig.width=15}
# make a versatile wide representation of the critical data
d_critical_predictorsXtarget <- d_critical_wide %>% 
  pivot_longer(
    cols = c(relevance, prior, competence, relevance_wUtt, competence_wUtt),
    names_to = "predictors",
    values_to = "predictor_response"
  ) %>%
  filter(predictors != "relevance_wUtt", predictors != "competence_wUtt")

d_critical_predictorsXtarget %>%
  ggplot(., aes(x = predictor_response, y = target)) +
  geom_point() +
  facet_wrap(title~predictors, ncol= 6) +
  xlab("Response to predictor questions") +
  ylab("Inference strength rating") +
  ggtitle("By-item Inference Strength Ratings vs. Predictor Ratings (left to right)")

@@

Plot inference strength ratings against predictor ratings, by-item, using by-item mean ratings across participants. Error bars indicate the standard deviation of the mean:
```{r, fig.height=10, fig.width=8}
d_critical_predictorsXtarget_summary <- d_critical_predictorsXtarget %>% 
  group_by(title, predictors, main_type) %>%
  summarise(mean_target = mean(target),
            mean_predictor = mean(predictor_response), 
            sd_target = sd(target),
            sd_predictor = sd(predictor_response))

d_critical_predictorsXtarget_summary %>%
  ggplot(., aes(x = mean_predictor, y = mean_target)) +
  geom_point() +
  geom_errorbar(aes(ymin = mean_target - sd_target, ymax = mean_target + sd_target), alpha = 0.7) +
  geom_errorbarh(aes(xmin = mean_predictor - sd_predictor, xmax = mean_predictor + sd_predictor), alpha = 0.7) +
  geom_text(aes(label=title), size = 2.5, nudge_y = 4) +
  geom_smooth(method = "lm") +
  ylab("Mean inference strength rating") +
  xlab("Mean predictor ratings") + 
  facet_wrap(main_type~predictors)
```

``` {r}

# correlation plot for "some"
GGally::ggpairs(
  filter(d_critical_wide, main_type == "some") %>%  
    select(prior, competence, relevance, target, competence_wUtt, relevance_wUtt) 
)

# correlation plot for "or"
GGally::ggpairs(
  filter(d_critical_wide, main_type == "xor") %>%  
    select(prior, competence, relevance, target, competence_wUtt, relevance_wUtt)
  )
```

## Explore z-scoring
Visual exploration of z-scoring each measure (rel / comp / pri / target / comp-wUtterance / rel-wUtterance) per participant, collapsing some and or together. 
```{r}
d_critical_zScored <- d_critical_long %>% group_by(submission_id, block_extended) %>%
  mutate(block_mean = mean(response),
         block_sd = sd(response),
         response_centered = (response - block_mean)/block_sd,
         # catch the cases where sd is 0 
         response_centered = ifelse(is.na(response_centered), 0, response_centered))

>>>>

---
title: "XOR-Some Prolific Pilot 3"
author: "Polina Tsvilodub"
date: "7/24/2021"
output: github_document
---

In this third pilot on Prolific we gather pilot data for the vignettes that were substantially updated based on the results from the [second pilot](https://github.com/magpie-ea/magpie-xor-experiment/blob/master/analysis/xor-some-prolific-pilot2.md). The used vignettes were selected based on manual comparison and comprise 38 out of 64 vignettes.

The structure of the experiment was as follows:
Participants read instructions, completed three example trials, and then completed 8 main blocks consisting of 4 xor and 4 some items. Each main block had the following structure: Participants read the background story, answered one comprehension question, then answered competence / relevance / prior questions in randomized order; then they read another 3 comprehension questions, after which the critical utterance was added below the background story. They answered the inference strength question. The second round of relevance/competence questions was removed for the pilot.

N=50 participants were recruited for this pilot and compensated 2 pounds/participant. 8 items (4 some + 4 xor) were sampled at random for each participant such that each participant saw one item in each condition (relevance X competence X prior = 8 unique conditions).

@@

```{r anonymization, include=FALSE}
#d_raw <- read_csv("~/Documents/Research/XOR/magpie-xor-experiment/data/pilots/results_63_xor-some-Prolific-pilot3_N50.csv")
# remove IDs
# d_raw %>% select(-prolific_id) %>% write_csv("../data/pilots/results_63_xor-some-Prolific-pilot3_N50.csv") 
```

## Checks & Exclusions
```{r data, results='hide', include=FALSE}
d <- read_csv("../data/pilots/results_63_xor-some-Prolific-pilot3_N50.csv")
```

@@

Check native languages. Participants not indicating English as (one of) their native language(s) are be excluded. We also compute some participant demographics.
```{r languages}
d %>% distinct(languages)
cat("Number of participants before excluding non-natives: ", d %>% distinct(submission_id) %>% count() %>% pull() ) 
# exclude non-natives if necessary
d_native <- d %>% 
 # filter(("en" | "En" | "bri") %in% languages)
  filter(grepl("[(en)(br)]", languages, ignore.case = T))

cat(" Number of participants after excluding non-natives: ", d_native %>% distinct(submission_id) %>% count() %>% pull() )

cat(" Mean age: ", d_native %>% pull(age) %>% mean(., na.rm = T) )
d_native  %>% count(gender) %>% mutate(n = n/72)
```

@@

```{r counts}
# check xor/some vs. trial type; there are more xor critical conditions because there are 2 prior questions
d_native %>% count(main_type, condition) 
# check experimental conditions
d_native %>% count(exp_condition)
# check xor/some vs. experimental condition
d_native %>% count(main_type, exp_condition)
# count items used
d_native %>% count(title)
```

@@

Investigate very fast completion times (we have none in this pilot):
```{r}
 d_native %>% mutate(timeSpent = round(timeSpent, 2)) %>%
  filter(timeSpent < 8) -> fast_completions

# check if their responses look somehow peculiar
#fast_completions %>%
#  ggplot(., aes(x = response)) +
#  geom_histogram()

@@

# get participants failing comprehension questions
d_test <- d_test %>%
  group_by(submission_id) %>%
  mutate(passed_filler_trial = case_when(test_condition == "true" ~ response >= 60,
                                   test_condition == "false" ~ response <= 40,
                                   test_condition == "uncertain" ~ response %in% (0:90)),
         mean_comprehension = mean(passed_filler_trial),
         passed_filler = mean_comprehension >= 0.8
         )

@@

```{r clean}
# get overall mean ratings / subject across comprehension and critical trials 
d_full_clean %>% group_by(submission_id) %>% summarise(mean_rating = mean(response)) %>% arrange(mean_rating)

# get mean ratings / subject in critical trials 
d_full_clean %>% filter(condition == "critical") %>% group_by(submission_id) %>% summarise(mean_rating = mean(response)) %>% arrange(mean_rating)

@@

@@

Create more extensive condition labels, including information about whether a rating was produced with or without the utterance given (no such condition in this pilot).

@@

Compute the average deviation from expected responses for each vignette, individually for each dimension (rel / comp / pri). Only half as many items deviate from expectations by >= 50 points compared to the last pilot.
```{r}
d_critical_long %>% 
  filter(block != "xor" & block != "some") %>%
  filter(block == class_condition) %>%
  mutate(expected_rating = prior_class * 100,
         response_deviation = abs(response - expected_rating),
         block = ifelse(block == "xor" | block == "some", "target", block)
         ) %>%
  group_by(title, block) %>%
  summarize(mean_deviation = mean(response_deviation)) %>%
  arrange(desc(mean_deviation), .by_group = T) %>%
  # get the largest deviations - over 50 
  filter(mean_deviation >= 50) -> d_critical_deviations

# d_critical_deviations %>% write_csv("../data/pilots/pilot3_byItem_rating_deviations.csv")

@@

# d_critical_sd %>% write_csv("../data/pilots/pilot3_byItem_rating_SDs.csv")
```

Get the vignettes and dimensions that were strongly deviating from expectations in the previous pilot and still did not improve:
```{r}
deviating_item_pilot2 <- read_csv("../data/pilots/pilot2_byItem_rating_deviations.csv") %>% select(-mean_deviation)
d_critical_deviations %>% select(-mean_deviation) %>% semi_join(., deviating_item_pilot2)
```

Visually check the spread of by-item ratings among subjects for each question:
```{r, echo=FALSE, fig.height=25, fig.width=10}
d_critical_long_grandMeans <- d_critical_long %>% 
  mutate(title = paste(title, exp_condition, sep = "_"),
         block = ifelse(block == "xor" | block == "some", "target", block)) %>%
  filter(block == class_condition | (block == "target" & class_condition == "relevance")) %>%
  group_by(main_type, block) %>% summarize(mean_resp = mean(response))

d_critical_long_expectedCats <- d_critical_long %>% 
  mutate(title = paste(title, exp_condition, sep = "_"),
         block = ifelse(block == "xor" | block == "some", "target", block)) %>%
  filter(block == class_condition | (block == "target" & class_condition == "relevance")) %>%
  group_by(main_type, block, title) %>% summarize(mean_resp_block = mean(response))

d_critical_long %>% 
  mutate(title = paste(title, exp_condition, sep = "_"),
         block = ifelse(block == "xor" | block == "some", "target", block)) %>%
  left_join(., d_critical_long_grandMeans, by = c("main_type", "block")) %>% 
  filter(block == class_condition | ((block == "target") & (class_condition == "relevance"))) -> d_long_plot_data

left_join(d_long_plot_data, d_critical_long_expectedCats, by = c("main_type", "block", "title")) %>% 
  mutate(expected_block = ifelse(block != "target", prior_class * 100, NA),
         block = factor(block, levels = c("relevance", "competence", "prior", "target"))) -> d_long_plot_full
d_long_plot_full %>%
  ggplot(., aes(x = block, y = response, color = block)) +
  geom_point(alpha = 0.7, position = position_jitter(width = 0.1)) +
  geom_hline(aes(yintercept = mean_resp), alpha = 0.7) +
  geom_point(aes(y = mean_resp), color = "black", shape = 2) +
  geom_point(aes(y = mean_resp_block, fill = block), color = "black") +
  geom_point(aes(y = expected_block), color="red", size = 2.5, na.rm = T) +
  ylab("Raw responses to respective questions") +
  facet_wrap(main_type~title, ncol = 4) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ggtitle("By-item by-question raw ratings\nBlack dots indicate by-block means within-item, lines and triangles indicate grand by-block means\nRed dots indicate expected rating")
```

Visually compare the updated vignettes (pilot 2 vs current pilot)
```{r, echo=FALSE}
# add column indicating which pilot it is
d_long_plot_full <- d_long_plot_full %>%
  mutate(pilot = "updated")

d_long_plot_pilot2 <- read_csv("../data/pilots/pilot2_byItem_plotting.csv") %>% 
  mutate(pilot = "old")
# only keep data for those items that were also tested in the current pilot
d_long_plot_pilot2 <- semi_join(d_long_plot_pilot2, d_long_plot_full, by = c("title"))

d_long_plot_both_pilots <- bind_rows(d_long_plot_full, d_long_plot_pilot2)
```

```{r, echo=FALSE, fig.width=8, fig.height=15}
# plot old and updated some items side-by-side
# the vignettes for which only an "updated" plot occurs are new items that were added instead of some other vignettes 
d_long_plot_both_pilots %>%
  filter(main_type == "some") %>%
  ggplot(., aes(x = block, y = response, color = block)) +
  geom_point(alpha = 0.7, position = position_jitter(width = 0.1)) +
  geom_hline(aes(yintercept = mean_resp), alpha = 0.7) +
  geom_point(aes(y = mean_resp), color = "black", shape = 2) +
  geom_point(aes(y = mean_resp_block, fill = block), color = "black") +
  geom_point(aes(y = expected_block), color="red", size = 2.5, na.rm = T) +
  ylab("Raw responses to respective questions") +
  facet_wrap(title~pilot, ncol = 4) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ggtitle("By-item by-question raw ratings, old vs. updated SOME vignettes\nBlack dots indicate by-block means within-item, lines and triangles indicate grand by-block means\nRed dots indicate expected rating")
```

```{r, echo=FALSE, fig.width=8, fig.height=15}
# plot old and updated xor items side-by-side 
d_long_plot_both_pilots %>%
  filter(main_type == "xor") %>%
  ggplot(., aes(x = block, y = response, color = block)) +
  geom_point(alpha = 0.7, position = position_jitter(width = 0.1)) +
  geom_hline(aes(yintercept = mean_resp), alpha = 0.7) +
  geom_point(aes(y = mean_resp), color = "black", shape = 2) +
  geom_point(aes(y = mean_resp_block, fill = block), color = "black") +
  geom_point(aes(y = expected_block), color="red", size = 2.5, na.rm = T) +
  ylab("Raw responses to respective questions") +
  facet_wrap(title~pilot, ncol = 4) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ggtitle("By-item by-question raw ratings, old vs. updated XOR vignettes\nBlack dots indicate by-block means within-item, lines and triangles indicate grand by-block means\nRed dots indicate expected rating")
```

@@

## Plots
```{r, echo=FALSE, eval=FALSE}
# Plot main rel / comp / prior questions by main condition and by prior classification of the item (x-axis), separated into with / without critical utterance (shape, color). The prior questions were only used without the utterance. 
d_critical_long %>% 
  filter(block != "xor" & block != "some") %>%
  filter(block == class_condition) %>%
  group_by(main_type, class_condition, w_utterance, prior_class) %>% 
  summarize(mean_response = mean(response)) -> d_critical_summary

d_critical_long %>% 
  filter(block != "xor" & block != "some") %>%
  filter(block == class_condition) %>%
  ggplot(., aes(x = as.factor(prior_class), y = response)) +
  geom_point(size = 2, alpha = 0.6, position = position_jitter(width = 0.1)) +
  geom_point(data = d_critical_summary, aes(x = as.factor(prior_class), y = mean_response), 
             color = "red", size = 3) +
  ylab("Responses to respective predictor question") +
  xlab("Anticipated categorization of the items (low / high)") +
  facet_wrap(main_type~class_condition) # get ratings from the respective trials only

```
First, compare the two prior ratings in the xor condition:
```{r, fig.height=3}
# check whether the ratings in the two prior questions are the same in xor 
d_critical_long %>% 
  filter(block != "xor" & block != "some") %>%
  filter(block == class_condition) %>%
  filter(block == "prior", main_type == "xor") %>%
  mutate(priorQ_nr = rep(c(1,2), nrow(.)/2),
         unique_ID = paste(submission_id, ID, sep="_")) -> d_xor_priors
d_xor_priors %>% group_by(priorQ_nr, prior_class) %>%
  summarise(mean = mean(response)) -> d_xor_priors_summary

d_xor_priors %>%
  ggplot(., aes(x = as.factor(priorQ_nr), y = response )) +
  geom_point(size = 2, alpha = 0.6) +
  geom_point(data = d_xor_priors_summary, aes(x = as.factor(priorQ_nr), y = mean), color = "red", size = 3) +
  geom_path(aes(group = "unique_ID"), alpha = 0.6) +
  ylab("Responses to prior questions") +
  xlab("First vs Second prior question for high vs low prior conditions") +
  facet_wrap(~as.factor(prior_class)) + # get ratings from the respective trials only 
  ggtitle("Xor prior question ratings.\nLines indicate the two responses provided by one participant for a specific item.")
```

```{r}
# check analytically how many subjects rate prior questions differently
d_xor_priors %>% select(-trial_number, -RT, -prompt) %>% 
  pivot_wider(names_from = "priorQ_nr", values_from = "response") %>%
  mutate(response_difference = abs(`1` - `2`),
         bigDifference = ifelse(response_difference > 15, T, F)) -> d_xor_priors_differences

# 19 responses with difference > 15 between first and second rating
d_xor_priors_differences %>% count(bigDifference)

@@

```{r, echo=FALSE, eval=FALSE}
# make a versatile wide representation of the critical data
d_critical_wide <- d_critical_long %>% 
  select(submission_id, title, main_type, block_extended, response) %>% 
  unique() %>% 
  pivot_wider(
    names_from = block_extended, 
    values_from = response, 
    values_fn = mean # getting means for double prior measurement in "xor"
  ) 
```
```{r, echo=FALSE, eval=FALSE}
# Plot inference strength ratings against raw predictor ratings from each participant across items:

d_critical_wide %>%
  ggplot(., aes(x = relevance, y = target)) +
  geom_point(size = 2, alpha = 0.7) +
  geom_smooth(method = "lm") +
  ylab("Inference strength ratings") +
  facet_wrap(~main_type, ncol = 1) -> p.rel

@@

gridExtra::grid.arrange(p.rel, p.comp, p.pri, ncol = 3) 
```
```{r, fig.width=10, fig.height=30, echo=FALSE, eval=FALSE}
# Plot mean ratings (across all participants) for each vignette (with its respective condition indicated in the facet title, in the order relevance/competence/prior), in each condition.
bar.width = 0.8
d_critical_long %>% 
  mutate(title = paste(title, exp_condition, sep = "_")) %>%
  filter(block == class_condition | block == "xor" | block == "some") %>%
  group_by(block_extended, title, w_utterance, main_type) %>%
  summarize(mean_rating = mean(response)) %>% 
  ggplot(., aes(x = block_extended, y = mean_rating)) +
  geom_col(alpha = 0.7, width = bar.width, position = position_dodge(width = bar.width)) +
  #geom_point(size = 2, alpha = 0.5, position = position_jitter(width = 0.1)) +
  ylab("Mean responses to respective questions") +
  xlab("Question type") +
  facet_wrap(main_type~title, ncol = 4) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ggtitle("By-item by-question mean ratings")
```
``` {r, echo=FALSE, eval=FALSE}

# correlation plot for "some"
GGally::ggpairs(
  filter(d_critical_wide, main_type == "some") %>%  
    select(prior, competence, relevance, target) 
)

# correlation plot for "or"
GGally::ggpairs(
  filter(d_critical_wide, main_type == "xor") %>%  
    select(prior, competence, relevance, target)
  )
```

As planned, all the primary visualizations and analyses below are conducted on data z-scored by-subject by-block.

Plot main rel / comp / prior questions by main condition and by prior classification of the item (x-axis): 
```{r}
d_critical_zScored <- d_critical_long %>% group_by(submission_id, block_extended) %>%
  mutate(block_mean = mean(response),
         block_sd = sd(response),
         response_centered = (response - block_mean)/block_sd,
         # catch the cases where sd is 0 
         response_centered = ifelse(is.na(response_centered), 0, response_centered))

d_critical_zScored %>% 
  filter(block != "xor" & block != "some") %>%
  filter(block == class_condition) %>%
  ggplot(., aes(x = as.factor(prior_class), y = response_centered, shape = w_utterance)) +
  geom_point(size = 2, alpha = 0.6, position = position_jitter(width = 0.1)) +
  geom_point(data = d_critical_zScore_summary, aes(x = as.factor(prior_class), y = mean_response), 
             color = "red", size = 3) +
  ylab("Responses to respective predictor questions") +
  xlab("Anticipated categorization of the items (low / high)") +
  facet_wrap(main_type~class_condition) + # get ratings from the respective trials only 
  ggtitle("Centered responses to by-predictor questions")
```
Plot inference ratings as a function of *anticipated* rating of the explanatory factor, similar to the paper: 
``` {r}
d_critical_zScored %>% 
  filter(block == "xor" | block == "some") %>%
  ggplot(., aes(x = prior_class, y = response_centered)) +
  geom_point(size = 2, alpha = 0.5) +
  geom_smooth(method="lm") +
  ylab("Inference strength ratings") +
  facet_wrap(block~class_condition) +
  xlab("Anticipated categorization of the items") +
  ggtitle("Inference strength ratings by-predictor")

@@

Plot the empirical predictor ratings against target inference ratings:
```{r}
d_critical_zScored_wide <- d_critical_zScored %>% 
  select(submission_id, title, main_type, block_extended, response_centered) %>% 
  unique() %>% 
  pivot_wider(
    names_from = block_extended, 
    values_from = response_centered, 
    values_fn = mean # getting means for double prior measurement in "xor"
  )
